---
title: "progress-report-nickdire214"
author: "Dominic DiRe"
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r packages}
library(rvest)
library(dplyr)
library(stringr)
library(readr)
library(Lahman)
library(ggplot2)
```


# Scrape the 2026 data

We read the HTML from the Baseball Reference 2026 HOF page and extract all tables.

```{r}
url <- "https://www.baseball-reference.com/awards/hof_2026.shtml"
supersecretsite <- read_html(url)
tables <- html_table(supersecretsite)
```

```{r}
head(tables)
```
Good comms on the row 1 being wack. The first row is the names of the columns.

```{r}
data <- tables[[1]] # pull
actual_col_names <- data[1, ] # set
colnames(data) <- actual_col_names # put
data <- data[-1, ] # scrap
head(data, 3) # money
```


# Cleaning data

Select the needed cols, convert num vars, take the "-X" out ,put "inducted" on 75%.

```{r}
hof_2026 <- data %>%
  select(Name, Votes, `%vote`) %>%
  mutate(
    Name = gsub("X-", "", Name),
    yearID = 2026,
    Votes = as.numeric(Votes),
    percent_vote = readr::parse_number(`%vote`),
    inducted = ifelse(percent_vote >= 75, "Y", "N")
  ) %>%
  rename(name = Name)

hof_2026
```


# Exporting cleaned (mine not used but this is if)

```{r}
# readr::write_csv(hof_2026, "HallOfFame_2026.csv")
```


# Compare with the Lahman

Filter Lahman to 2026 and compare counts

```{r}
lahman_2026 <- HallOfFame %>%
  filter(yearID == 2026)

hof_2026 %>%
  count(inducted)

lahman_2026 %>%
  count(votedBy, inducted)
```



# Conclusion

The scraped baseball dataset and the Lahman version talk about the same thing but are not presented the same. You kinda expect differences with this kinda thing, but I was surprised the Lahman data only included people voted by the Contemporary Baseball Era

The scraped dataset overall shows every canidate who received one or more hall of fame vote. The Lahman dataset has a "votedBy" column which is not in the scraped data at all. Which like I said above inclueded one group of voting. The scraped data had 25 canidates and 2 above 75% (labeled inducted), whereas Lahman had 8 canidates and 1 inducted. 
